Converting LaTeX R Journal Articles into RJ-web-articles

### 2023-06-24 texor.Rmd
notes on more to include in intro
Jotted down a part about rebib

---
title: "Converting LaTeX R Journal Articles into RJ-web-articles"
date: "2022-11-09"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Abhishek Ulayil
    affiliation: Institute of Actuaries of India
    address:
    - Mumbai, India
    orcid: 0009-0000-6935-8690
    email: perricoq@outlook.com
  - name: Mitch O'Hara-Wild
    affiliation: Monash University
    address:
    - Melbourne, Australia
    orcid: 0000-0001-6729-7695
    email: mail@mitchelloharawild.com
  - name: Christophe Dervieux
    affiliation: Posit PBC
    address:
    - Paris, France
    orcid: 0000-0003-4474-2498
    email: christophe.dervieux@gmail.com
  - name: Heather Turner
    affiliation: University of Warwick
    address:
    - Newport, United Kingdom
    orcid: 0000-0002-1256-3375
    email: ht@heatherturner.net
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Melbourne, Australia
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu
type: package
output: 
  rjtools::rjournal_web_article:
    web_only: yes
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

@@

@@

The Texor package uses these filters heavily, even for simple tasks like choosing
which supporting figures to copy. A lua filter constructs a list of image paths, 
writes it to an intermediate file and feeds them to the Texor package.

@@

LaTeX is very descriptive and allows unfettered levels of customization to authors,
which poses an issue during conversions. RMarkdown is quite a bit restrictive by 
syntax though you can always embed HTML to expand the feature set. Although 
pandoc does most of the heavy lifting, there are some instances where even 
filters cannot assist. A good demonstration of the issue in hand is code 
environments.

A limitation of pandoc is that it only works with a "verbatim" environment. 
However, in "RJournal",  we have multiple custom code environments which are not
supported or defined due to lower-level tex limitations in pandoc. Hence the code 
environments used to be eliminated during the conversion process, including a Lua
filter to find these elements and redefine them as code environments would also 
fail because pandoc would have already tried to process the code inside an unknown 
code environment, leading to missing code elements lost in the conversion.

@@

This was one of the earliest and most basic Lua filters, and it was part of the 
texor package:

@@

# texor package converting LaTeX to Rmarkdown

## Also mention the hard parts
## Lua Filters with pandoc
A lot of details about different components of texor

# rebib package working out bibliography

>>>>

---
title: "Converting LaTeX R Journal Articles into RJ-web-articles"
date: "2022-11-09"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Abhishek Ulayil
    affiliation: Student, Institute of Actuaries of India
    address:
    - Mumbai, India
    orcid: 0009-0000-6935-8690
    email: perricoq@outlook.com
  - name: Mitch O'Hara-Wild
    affiliation: Monash University
    address:
    - Melbourne, Australia
    orcid: 0000-0001-6729-7695
    email: mail@mitchelloharawild.com
  - name: Christophe Dervieux
    affiliation: Posit PBC
    address:
    - Paris, France
    orcid: 0000-0003-4474-2498
    email: christophe.dervieux@gmail.com
  - name: Heather Turner
    affiliation: University of Warwick
    address:
    - Newport, United Kingdom
    orcid: 0000-0002-1256-3375
    email: ht@heatherturner.net
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Melbourne, Australia
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu
type: package
output: 
  rjtools::rjournal_web_article:
    web_only: yes
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

@@

- R Journal latex template and formatting
- Quirks of the legacy articles

@@

The Texor package uses these filters heavily, even for simple tasks like choosing
which supporting figures to copy. A Lua filter constructs a list of image paths, 
stores them momentarily and the package uses this data to copy images.

@@

LaTeX is very descriptive and allows unfettered levels of customization to authors,
which poses an issue during conversions. RMarkdown is quite restrictive by 
syntax though you can always embed HTML to expand the feature set. Although 
pandoc does most of the heavy lifting, there are some instances where even 
filters cannot assist. A good demonstration of the issue in hand is code 
environments.

A limitation of pandoc is that it only works with a "verbatim" environment. 
However, in "RJournal",  we have multiple custom code environments which are not
supported or defined due to lower-level tex limitations in pandoc. So the code 
environments are eliminated during the conversion process. Including a Lua filter
to find these elements and redefine them as verbatim environments would also be 
ineffective. If the code environment is not verbatim, then pandoc will also try 
to process the actual code content as LaTeX commands and will lose out on 
details in some cases.

@@

This was one of the earliest and most basic Lua filters, and it was part of the 
\CRANpkg{texor} package:

@@

# Texor package converting LaTeX to Rmarkdown

## Also mention the hard parts
## Lua Filters with pandoc
A lot of details about different components of \CRANpkg{texor}

# Rebib is in charge of the bibliography

The \CRANpkg{rebib} package addresses the issue with LaTeX articles using 
built-in bibliography options with or without BibTeX files. While this works 
well with LaTeX, it won't work with Rmarkdown. Initially the goal was to use 
external software like Biber to convert the embedded bibliography to BibTeX. 
However the integration of external software was not viable, hence the 
experimental idea of a bibliography parser gained momentum.

It was initially a part of the \CRANpkg{texor} package, as those functions grew 
in features and became more involved; at that point, it made sense to move those
functions as a separate package. Initially, there were some reservations about 
the usage of rebib and its stability with various formats. However,
the package has improved over time and proven to be a good performer.

Rebib's best feature is its ability to merge bibliographies from both the 
LaTeX and BibTeX files. \CRANpkg{rebib} saves authors from manually 
converting to BibTeX format.
